---
title: "assignment3"
author: "Anna Calle"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(sensitivity)
library(pse)
```

Adjust your almond model to output ONLY the mean almond yield anomoly IF the users sets parameter (e.g mean_only = TRUE))

Perform a sensitivity analysis of how mean anomaly varies ALL of the parameters used in the yield model  

Assume parameters are normally distributed with standard deviation of 20% mean value

Rank the parameters in term of their sensitivity

Graph uncertainty in mean yield anomaly across all parameter uncertainty (boxplot and cummulative distribution of the output).

Repeat using the LHS and Sobel methods

Repeat using twice as many parameter sets as you did in your first sensitivity analysis - and look at how this changes the sensitivity results

Submit R markdown and short write up describing what you learned from the sensitivity analysis. Please submit your markdown as an .html or PDF. 

```{r}
# read in the input data
clim <- read.table("clim.txt",  sep=" ", header=T)

# source function
source("almond_yield.R")

almond_yield(mean_only = TRUE)
```


LHS
```{r}
# names of our parameters
factors = c("a", "b", "c", "d", "e")

# type of distributions they arise from
q = c("qnorm", "qnorm", "qnorm", "qnorm", "qnorm")

# parameters mean
a = -0.015
b = -0.0046
c = -0.07
d = 0.0043
e = 0.28

# parameters sd
sd_a = 0.2 * a
sd_b = 0.2 * b
sd_c = 0.2 * c
sd_d = 0.2 * d
sd_e = 0.2 * e

# parameters for those distributions
q.arg = list(list(mean=a, sd=0.2),
             list(mean=b, sd=0.2),
             list(mean=c, sd=0.2),
             list(mean=d, sd=0.2),
             list(mean=e, sd=0.2))

nsets=200
sens_YA = LHS(NULL,factors,nsets,q,q.arg, nboot=500)
# save parameter values
sens.pars = get.data(sens_YA)

head(sens.pars)

ggplot(sens.pars, aes(a))+geom_density()
ggplot(sens.pars, aes(x=a, y=b))+geom_density_2d()

```

```{r runsen, echo=TRUE, eval=TRUE}

# run the model
sens_res = almond_yield(clim_data = clim, mean_only = TRUE, a=sens.pars$a, b=sens.pars$b, c=sens.pars$c, d=sens.pars$d, e=sens.pars$e)

sens_YA = pse::tell(sens_YA, sens_res, res.names="YA")
plotscatter(sens_YA)

# note it can take standard R plotting parameters
plotscatter(sens_YA, col="darkgreen", cex=5, ylab="Yield Anomaly (ton/acre)")

# whats is the range of results

plotecdf(sens_YA, col="red", lwd=5, xlab="Yield Anomaly (ton/acre)")

# we can also plot partial correlation coefficients
plotprcc(sens_YA)
# and we can look at the actual values
sens_YA$prcc
# note because this is a linear model correlations are high for both parameters - and similar

```






